---
title: Network analysis (R)
output: html_document
---

# Network analysis

Network analysis focuses on seeing the world via nodes and ties between nodes.
It can be used for various purposes, ranging from [textual data](https://sicss.io/2020/materials/day3-text-analysis/text-networks/rmarkdown/Text_Networks.html) to analysis of [sexual relationships in a high school](https://people.duke.edu/~jmoody77/chains.pdf).

Now we learn how to
* read data in as a network
* visualise network
* compute various network characteristics

For this, we use the package [igraph](https://igraph.org/r/).

[Narrated code walkthrough](https://www.youtube.com/watch?v=lXC2egj8lmY)

```{r}
library(igraph)
```

```{r}
data <- read.csv('data/org_x_collaboration.csv', header = F )
matrix <- as.matrix( data )

network <- graph_from_adjacency_matrix( matrix, 'directed' )
```

## Computing network characteristics

There are various measurements which can be executed both at the level of individual nodes as well as at the level of the whole network.

```{r}
degree( network )
summary( degree( network ) )
hist( degree( network ) )
```

## Exercises

* Instead of this ugly histogram, use `ggplot` to draw a nice histogram.
* Instead of degree use [betweness](https://igraph.org/r/doc/betweenness.html) metrics.
* Do degree and betweness correlate?
* What is the diameter of the network?

## Plotting networks

Often network analysis is done through drawing the network.
Here we learn how to twist the drawing by changing node attributes, tie attributes as well as the [layout of the graph](https://igraph.org/r/doc/layout_.html).
See [documentation](https://igraph.org/r/doc/plot.common.html) for various alternatives you can use.

```{r}
plot( network )
```

```{r}
V(network)$color <- 'white'
V(network)$size <- degree(network)

E(network)$color <- 'green'

l = layout_in_circle( network )

plot( network, layout = l )
```

