---
title: An R Markdown document converted from "Simulation - system dynamics (R).ipynb"
output: html_document
---

# Simulations: system dynamics

System dynamics models work on stocks and flows between the stocks.
To get an idea about this, see how [MONIAC](https://www.youtube.com/watch?v=rVOhYROKeu4) works.
See the documentation of [deSolve](https://cran.r-project.org/web/packages/deSolve/index.html).

```{r}
library(deSolve)
library(tidyverse)
```

```{r}
## time starts from 0 and continues until 20

time <- seq(0, 20, by=1)

## we have three stocks (people healthy, people sick and people killed)

stocks <- c( healthy = 900, sick = 100, killed = 0 )

## model has parameters, such as infection rate amd kill rate

auxs <- c( infection = .05, dead = 0.05 )
```

```{r}
## here we define how each step is executed

model <- function(time, stocks, auxs){
  with(as.list(c(stocks, auxs)), {
      
    # we can modify parameters over time
    # if( time >= 10 ) {
    #     infection <- .00
    # }

    new_people_sick <- sick * infection
    new_dead_people <- sick * dead
      
    healthy_change <- - new_people_sick
    dead_change <- new_dead_people
    sick_change <- - new_dead_people + new_people_sick

    ## we update all three stocks and parameters
    return (
        list(
            c( healthy_change, sick_change, dead_change ),
            auxs
        )
    )
  }) 
}
```

```{r}
## this runs the model
data <- data.frame( dede( y = stocks, times = time, func = model, parms = auxs, method = "lsodar"))
head( data )
tail( data )
```

```{r}
df <- data %>% 
  pivot_longer(-time, names_to = 'variable')


ggplot(df, aes(time, value, color = variable))+
  geom_line(size =1.25)+
  theme_minimal()
```

## Exercises

1. Based on the model equation, draw a picture about the model (similar to Figure 6.2)
1. Change the infection rate to 1.10 and dead rate to 0.05 and plot the results.
1. Increase the time span by 10 more days.
1. Examine how a stronger intervention (decreasing the infection rate more) impacts to number of patients and deadths. What do you see?
1. Change the time of the intervention, what changes can you detect?
1. Examine the timespan to 300 days. Why does the model break? Fix it so that it does not!

